---
title: |
  <div class="tight-title">
    <div>ECOM30004/90004</div>
    <div>Time Series Analysis and Forecasting</div>
     <div class="spacer"></div> <!-- blank line -->
    <div class="bold">Tutorial 7</div>

    <div class="spacer"></div> <!-- blank line -->

    <div class="author">Tutor: Pengyun (Kevin) Chen</div>
    <div class="author">University of Melbourne</div>
    <div class="author">Semester 2, 2026</div>
  </div>

  
format:
  revealjs:
    theme: simple
    self-contained: true
    chalkboard: false
    width: 1300
    height: 950
    footer: 'Access Slides: `pengyunc.github.io/TSAF_Week7`'
    title-slide-attributes:
      data-background-color: "#000F46"
    slide-number: true
    center-title-slide: true
    toc: false
    css: styles.css
    callout-appearance: simple
    background-transition: fade
    callout-icon: true
    transition: slide
    scrollable: true
    code-block-height: 800px
    code-line-numbers: true
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-svg.js"
    html-math-method: mathjax
    execute:
      echo: true
      eval: true
      enabled: true
---

## Notice

::: {.callout-warning title="Conditions of Use"}

- This document is intended for personal use and is NOT meant to be shared on the internet (e.g. Studocu, ThinkSwap, StudentVIP, Coursehero, etc.)
- Do not share this document with anyone, even if they are enrolled in ECOM30004/90004.
- Official course materials, such as lecture slides, tutorial sheets and prescribed readings should always be your primary point of reference.

:::
<div style="height: 0.2em;"></div>
 - Slides accessible here: `pengyunc.github.io/TSAF_Week7`
 - Download HTML versions here: [Dropbox Link](https://www.dropbox.com/scl/fo/aoxoud5ex0j42hsrod3dg/ADblAt_jJ9hpESZME9FDcWc?rlkey=0s0oyf1y3uokzvlhbzxe2yy3g&st=kx2yupzr&dl=0)

## Recap of Tutorial 6

<blue>In Tutorial 6, the ADF test indicated that the $\log(CPI_{t})$ series contains a unit root. </blue>

:::{.incremental}

- This means we cannot just remove the trend and fit an ARMA model.

  - <teal>$\to$ We need to first difference the $\log(CPI_{t})$ series to render it stationary.</teal>

- We then ran the ADF test with drift on $\Delta\log(CPI_{t})$ and the null hypothesis of a unit root was rejected at 5\% level. 

  - <oblue>$\to$ $\Delta\log(CPI_{t})$ can be modeled as an ARMA process.</oblue>
  - So we proceeded with selecting the optimal ARMA($p,q$) order for $\Delta\log(CPI_{t})$ using a grid search of $p=0,\dots,4$ and $q=0,\dots,4$.

- <blue>**So what is the purpose of ADF test here?**</blue>

  - To assist with selecting $d=0$ or $d=1$ in an ARIMA model.
  
  - We may wonder: Can we let an information criterion (AIC or AICc) select $d$ automatically, alongside $p$ and $q$?
  
- <blue>**This tutorial:** Maybe we can replace that decision with AIC, just like we pick $p$ and $q$?</blue> <green>$\to$ we will examine this idea with a simulation</green>
  
:::


## This tutorial {auto-animate="true"}
<blue>In this tutorial we will use simulation to explore the possible usefulness of the AIC for deciding whether to difference a time series or not, in comparison to the conventional ADF tests.</blue>

<div style="height: 0.2em;"></div>
:::{.callout-note title = "Simulation Setting"}

:::{.incremental}

- Use a sample size of $n=100$.
- Fit an AR model with fixed order $p=2$ and with intercept only:
$$\mathrm{ARIMA}(2,0,0):\qquad Y_{t}=\beta_{0}+\phi_{1}Y_{t-1}+\phi_{2}Y_{t-2}+U_{t}$$
- A restricted version of this model is to take a first difference of $Y_{t}$:
$$\mathrm{ARIMA}(1,1,0):\qquad\Delta Y_{t}=\beta_{0}+\psi_{1}\Delta Y_{t-1}+U_{t}$$
- <blue>Selection Criteria: </blue>Select ARIMA(1,1,0) if
$\operatorname{AIC}_{\operatorname{ARIMA(1,1,0)}}<\operatorname{AIC}_{\operatorname{ARIMA(2,0,0)}}$.

- We will compare the ADF and AIC approaches in two ways:
  i. proportion of “correct” choice of differencing
  ii. resulting one-step-ahead forecasting performance.
  
:::
:::




## Simulation: Data Generating Process {auto-animate="true"}

:::{.callout-note title="Data Generating Process"}
The data generating process for the simulation will be

$$Y_{t}=\phi Y_{t-1}+U_{t},\qquad U_{t}\sim\text{i.i.d. N}(0,1)$$
with $n=100$ and $\phi_{1}=1,0.95,0.9,0.85,0.8,0.7,0.6,0.5$.
:::

<div style="height: 0.5em;"></div>

- For setting up the code, we select $\phi=0.9$ (without loss of generality), but we can change it later.

- The true underlying process is actually AR(1), although we will fit AR(2).

- When $\phi_{1}\ne 1$, no differencing is needed for $Y_{t}$ <teal>$\to$ the correct choice is AR(2)</teal>

- When $\phi_{1}=1$, $Y_{t}$ is a unit-root process <purple>$\to$ the correct choice is ARIMA(1,1,0)</purple>

## Initialisation {auto-animate="true"}

Load libraries:
```{r}
library(forecast)
library(urca)
```
<div style="height: 0.5em;"></div>

Set parameters:
```{r}
n <- 100        # Sample size
phi1 <- 0.9     # True value of phi1
p <- 2          # Lag length of AR model
t <- (p+1):n    

reps <- 1000    # Number of repetitions
```
<div style="height: 0.5em;"></div>

Initialise Container Matrices:
```{r}
# For storing model selection outcomes by ADF and AIC
ur <- matrix(nrow=reps, ncol=2)
colnames(ur) <- c("ADF", "AIC")

# Matrix of forecast errors for each replication
ForecastErrors <- matrix(nrow=reps, ncol=4)
colnames(ForecastErrors) <- c("ARIMA110", "ARIMA200", "ADF", "AIC")
```


## Initialisation {auto-animate="true"}
Initialise Container Matrices:
```{r}
# For storing model selection outcomes by ADF and AIC
ur <- matrix(nrow=reps, ncol=2)
colnames(ur) <- c("ADF", "AIC")

# Matrix of forecast errors for each replication
ForecastErrors <- matrix(nrow=reps, ncol=4)
colnames(ForecastErrors) <- c("ARIMA110", "ARIMA200", "ADF", "AIC")
```
<div class="horizontal-line"></div>
```{r}
print(ur)        # This should be a matrix of 0's and 1's
```

::: {.absolute top="560px" left="480px" style="font-size: 36px;"}
<blue>Store model selection results of `ADF` (Column 1) and `AIC` (Column 2):</blue>
:::

::: {.absolute top="720px" left="500px" style="font-size: 36px;"}
unit root/differencing chosen $\to$ record `1` 
:::

::: {.absolute top="670px" left="500px" style="font-size: 36px;"}
no differencing chosen $\to$ record `0` 
:::

## Initialisation {auto-animate="true"}
Initialise Container Matrices:
```{r}
# For storing model selection outcomes by ADF and AIC
ur <- matrix(nrow=reps, ncol=2)
colnames(ur) <- c("ADF", "AIC")

# Matrix of forecast errors for each replication
ForecastErrors <- matrix(nrow=reps, ncol=4)
colnames(ForecastErrors) <- c("ARIMA110", "ARIMA200", "ADF", "AIC")
```
<div class="horizontal-line"></div>
```{r}
print(ForecastErrors)      # This should be a matrix of forecast errors
```

::: {.absolute top="560px" left="600px" style="font-size: 36px;"}
<blue>Store (one-step-ahead) forecast errors of:</blue>
:::

::: {.absolute top="620px" left="600px" style="font-size: 36px;"}
`ARIMA110`: ARIMA(1,1,0) model
:::

::: {.absolute top="680px" left="600px" style="font-size: 36px;"}
`ARIMA200`: ARIMA(2,0,0) model
:::

::: {.absolute top="740px" left="600px" style="font-size: 36px;"}
`ADF`: Model selected by ADF test
:::

::: {.absolute top="800px" left="600px" style="font-size: 36px;"}
`AIC`: Model selected by AICc
:::


## Simulation Code {auto-animate="true"}
<blue>Code structure:</blue>
```{r eval=FALSE}
set.seed(42)
for (r in 1:reps){

  # a. Generate Y

  # b. Estimate ARIMA(2,0,0) and ARIMA(1,1,0) models for Y

  # c. Compute and store 1-step-ahead forecast errors for each model
  #    in columns 1 and 2 of ForecastErrors

  # d. ADF test on Y, store decision in column 1 of ur
  #    Store forecast error of ADF-chosen model in col 3 of ForecastErrors

  # e. AIC choice between ARIMA(2,0,0) and ARIMA(1,1,0)
  #    Store decision in column 2 of ur
  #    Store forecast error of AIC-chosen model in col 4 of ForecastErrors

}
```



## Simulation Code {auto-animate="true"}
<blue>**Part a:** Generate data from an AR(1) model: $Y_t=\phi_{1}Y_{t-1}+U_{t}$.</blue>
```{r eval=FALSE}
  # a. Generate Y
  
  if (phi1==1){
    Y <- arima.sim(n=n+1, model=list(order=c(0,1,0)))[-1]
  } else {
    Y <- arima.sim(n=n+1, model=list(ar=phi1))
  }
  
  # Keep aside Y[n+1] (last element) for forecast evaluation
  Yf <- Y[n+1]
  # Use Y1,...,Yn (first n elements) as the estimation sample
  Y <- Y[1:n]
```
<div style="height: 0.5em;"></div>
When $\phi_{1}=1$, the model becomes a random walk (`[-1]` removes the first entry):
```{r}
Y <- arima.sim(n=n+1, model = list(order = c(0,1,0)))
print(Y)
```


## Simulation Code {auto-animate="true"}
<blue>**Part b:** Estimate ARIMA(2,0,0) and ARIMA(1,1,0) models for $Y_{t}$</blue>
```{r eval=FALSE}
  # b. Estimate ARIMA(2,0,0) and ARIMA(1,1,0) models for Y

  # Estimation of AR(2) model
  ARIMA200 <- lm(Y[t]~Y[t-1]+Y[t-2])
  
  # Estimation of ARIMA(1,1,0) model
  DY <- Y[t]-Y[t-1]
  DY1 <- Y[t-1]-Y[t-2]
  ARIMA110 <- lm(DY[t]~DY[t-1])
```
:::{.callout-tip title="Notes"}
- Recall we set `t = (p+1):n`, and `p = 2` is the AR order.
- <blue>**Why do we use `lm` instead of `Arima` this time?**</blue>
  - This is because when $\phi_{1}=1$, the series is not stationary.
  - `Arima` command enforces a stationarity check and will return an error in this case.
  - Beneath the surface `Arima` is attempting to do “maximum likelihood” estimation, which can be tricky in non-stationary settings, while `lm` is doing least squares which continues to be workable.
:::



## Simulation Code {auto-animate="true"}
<blue>**Part c:** Compute one step ahead forecast errors</blue>
```{r eval=FALSE}
  # c. Compute and store 1-step-ahead forecast errors for each model
  #    in columns 1 and 2 of ForecastErrors

  b1 <- ARIMA110$coefficients
  ForecastErrors[r,1] <- Yf - (Y[n]+b1[1]+b1[2]*DY[length(DY)])

  b2 <- ARIMA200$coefficients
  ForecastErrors[r,2] <- Yf - (b2[1]+b2[2]*Y[n]+b2[3]*Y[n-1])
```

:::{.callout-tip title="Notes"}
- `Yf` is the actual 1-step-ahead observation ($Y_{n+1}$)
- Forecast errors are computed as:
$$\hat{U}_{n+1}=Y_{n+1}-\hat{\mathbb{E}}[Y_{n+1}\mid\mathscr{Y}_{n}]$$
where the forecasting equations are:
$$\begin{aligned}\text{ARIMA(1,1,0):}&& \hat{\mathbb{E}}[Y_{n+1}\mid\mathscr{Y}_{n}]&=Y_{n}+\color{blue}{\widehat{\Delta Y}_{n+1}}\,\color{black}{=Y_{n}\,+}\,\color{blue}{(\hat{\beta}_{0}+\hat{\psi}_{1}\Delta Y_{n})}\\
\text{ARIMA(2,0,0):}&& \hat{\mathbb{E}}[Y_{n+1}\mid\mathscr{Y}_{n}]&=\hat{\beta}_{0}+\hat{\phi}_{1}Y_{n}+\hat{\phi}_{2}Y_{n-1}
\end{aligned}$$
:::



## Simulation Code {auto-animate="true"}
<blue>**Part d:** Perform the ADF test to check for differencing</blue>
```{r eval=FALSE}
  # d. ADF test on Y, store decision in column 1 of ur

  # ADF test with drift, lags=1 (include one augmenting term)
  adf_tstat = ur.df(Y, type="drift", lags=1)@teststat[1]
  ur[r,1] <- 1*(punitroot(adf_tstat, N=n, trend="c", statistic="t")>0.05)
  
  # Store forecast error of ADF-chosen model in col 3 of ForecastErrors
  
  # If ADF test selects unit root use ARIMA110 forecast error:
  if (ur[r,1]==1){
    ForecastErrors[r,3] <- ForecastErrors[r,1]  
  } else { 
    # otherwise use ARIMA200 forecast error:
    ForecastErrors[r,3] <- ForecastErrors[r,2]  
  }
```
<div style="height: 0.5em;"></div>

Recall the hypotheses of the ADF test are:
\begin{align}
H_{0}:&\;\varphi=0& &\text{($Y_{t}$ has a unit root)}\\
H_{1}:&\;\varphi<0& &\text{($Y_{t}$ does not have a unit root)}
\end{align}
thus if $p>0.05 \implies H_{0}$ is maintained, ADF suggests ARIMA(1,1,0) <blue>$\to$ **record 1**</blue>


## Simulation Code {auto-animate="true"}
<blue>**Part e:** Use AIC to check for differencing</blue>
```{r eval=FALSE}
  # e. AIC choice between ARIMA(2,0,0) and ARIMA(1,1,0)

  # Store decision in column 2 of ur
  ur[r,2] <- 1*(AIC(ARIMA110)<AIC(ARIMA200))

  # Store forecast error of AIC-chosen model in col 4 of ForecastErrors
  if (ur[r,2]==1){
    # If AIC selects unit root use ARIMA110 forecast error:
    ForecastErrors[r,4] <- ForecastErrors[r,1]  
  } else {
    # otherwise use ARIMA200 forecast error:
    ForecastErrors[r,4] <- ForecastErrors[r,2]  
  }
```
<div class="horizontal-line"></div>
Recall that when AIC is used, we select ARIMA(1,1,0) if
$$\operatorname{AIC}_{\operatorname{ARIMA(1,1,0)}}<\operatorname{AIC}_{\operatorname{ARIMA(2,0,0)}}$$
if this is satisfied, store `ur[r,2]=1`, and store the forecast error of ARIMA(1,1,0) under `ForecastErrors[r,4]`.


## Simulation Results

To make the results reproducible, we set a random seed at `42`:
```{r eval=FALSE}
set.seed(42)    
```
<div class="horizontal-line"></div>
```{r echo=FALSE}
set.seed(42)
for (r in 1:reps){
  
  # Simulate time series of length n+1
  #   ARIMA(0,1,0) if phi1=1, ARIMA(1,0,0) otherwise
  if (phi1==1){
    Y <- arima.sim(n=n+1, model=list(order=c(0,1,0)))[-1]
  } else {
    Y <- arima.sim(n=n+1, model=list(ar=phi1))
  }
  
  # Keep aside Y[n+1] for forecast evaluation
  Yf <- Y[n+1]
  # Use Y1,...,Yn as the estimation sample
  Y <- Y[1:n]
  
  # Estimation of AR(2) model
  ARIMA200 <- lm(Y[t]~Y[t-1]+Y[t-2])
  
  # Estimation of ARIMA(1,1,0) model
  DY <- Y[t]-Y[t-1]
  DY1 <- Y[t-1]-Y[t-2]
  ARIMA110 <- lm(DY[t]~DY[t-1])
  
  
  # Compute one step ahead forecast errors
  b1 <- ARIMA110$coefficients
  ForecastErrors[r,1] <- Yf-(Y[n]+b1[1]+b1[2]*DY[length(DY)])
  
  b2 <- ARIMA200$coefficients
  ForecastErrors[r,2] <- Yf-(b2[1]+b2[2]*Y[n]+b2[3]*Y[n-1])
  
  
  # ADF test with drift, lags=1 (for p=2 model)
  # p>0.05 => H0:unit root not rejected, i.e. differencing indicated
  ur[r,1] <- 1*(punitroot(ur.df(Y, type="drift", lags=1)@teststat[1],
                          N=n, trend="c", statistic="t")>0.05)
  
  # If ADF test selects unit root use ARIMA110 forecast error:
  if (ur[r,1]==1){
    ForecastErrors[r,3] <- ForecastErrors[r,1]  
  } else { 
    # otherwise use ARIMA200 forecast error:
    ForecastErrors[r,3] <- ForecastErrors[r,2]  
  }
  
  # Use AIC to select between AR(2) and ARIMA(1,1,0)
  ur[r,2] <- 1*(AIC(ARIMA110)<AIC(ARIMA200))
  
  # If AIC selects unit root use ARIMA110 forecast error:
  if (ur[r,2]==1){
    ForecastErrors[r,4] <- ForecastErrors[r,1]  
  } else {
    # otherwise use ARIMA200 forecast error:
    ForecastErrors[r,4] <- ForecastErrors[r,2]  
  }
  
}
```

Running the completed code produces the following results:
```{r}
head(ur)
```
<div style="height: 0.5em;"></div>
```{r}
head(ForecastErrors)
```


## Simulation Results
:::{.callout-note}
<blue>We will compare the ADF and AIC approaches in two ways:</blue>

i. proportion of “correct” choice of differencing
ii. resulting one-step-ahead forecasting performance.
:::
<div style="height: 0.25em;"></div>
**When $\phi_{1}=0.9$:**

<blue>Calculate the selection proportions for the ADF and AIC:</blue>
```{r}
# Selection proportions:
UnitRootSelected <- apply(ur,2,mean)*100
print(UnitRootSelected)
```
<div style="height: 0.5em;"></div>
<blue>RMSE’s of (i) always using ARIMA(1,1,0), (ii) always using ARIMA(2,0,0), (iii) choosing between them with ADF test, (iv) choosing between them with AIC:</blue>
```{r}
RMSE <- sqrt(apply(ForecastErrors^2,2,mean))
print(round(RMSE, 5))
```


## Simulation: Looping over all $\phi_{1}$ values  {auto-animate="true"}
Modify the code to consider all of $\phi_{1}=1,0.95,\dots,0.5$:
```{r}
n <- 100        # Sample size
p <- 2          # Lag length of AR model
t <- (p+1):n
reps <- 1000    # Number of repetitions

# True value of phi1
phi1_values <- c(1, 0.95, 0.9, 0.85, 0.8, 0.7, 0.6, 0.5)
```
<div style="height: 1em;"></div>
Set up two extra storage matrices for model selection results and RMSEs:
```{r}
# Stores model selection results for each phi1 (should be 8x2)
my_ur = matrix(nrow=length(phi1_values), ncol=2)
colnames(my_ur) <- c("ADF", "AIC")
rownames(my_ur) <- paste0("phi1=", phi1_values)

# Stores RMSE results for each phi1 (should be 8x4)
my_RMSE = matrix(nrow=length(phi1_values), ncol=4)
colnames(my_RMSE) <- c("ARIMA110", "ARIMA200", "ADF", "AIC")
rownames(my_RMSE) <- paste0("phi1=", phi1_values)
```

## Simulation: Looping over all $\phi_{1}$ values  {auto-animate="true"}
Next, add an extra layer of `for` loop for `phi1` ($\phi_{1}$) values:
```{r eval=FALSE}
for (i in 1:length(phi1_values)){
  
  phi1 = phi1_values[i]
  
  ur <- matrix(nrow=reps, ncol=2)
  colnames(ur) <- c("ADF", "AIC")
  
  ForecastErrors <- matrix(nrow=reps, ncol=4)
  colnames(ForecastErrors) <- c("ARIMA110", "ARIMA200", "ADF", "AIC")
  
  set.seed(42)
  for (r in 1:reps){
    
    # steps a-e code goes here...
    
  }
  
  my_ur[i,] <- apply(ur, 2, mean)*100
  my_RMSE[i,] <- sqrt(apply(ForecastErrors^2, 2, mean))
  
}
```


## Results {auto-animate="true"}
```{r echo=FALSE}
my_ur = matrix(nrow=length(phi1_values), ncol=2)
colnames(my_ur) <- c("ADF", "AIC")
rownames(my_ur) <- paste0("phi1=", phi1_values)

my_RMSE = matrix(nrow=length(phi1_values), ncol=4)
colnames(my_RMSE) <- c("ARIMA110", "ARIMA200", "ADF", "AIC")
rownames(my_RMSE) <- paste0("phi1=", phi1_values)

for (i in 1:length(phi1_values)){
  
  phi1 = phi1_values[i]
  
  ur <- matrix(nrow=reps, ncol=2)
  colnames(ur) <- c("ADF", "AIC")
  
  ForecastErrors <- matrix(nrow=reps, ncol=4)
  colnames(ForecastErrors) <- c("ARIMA110", "ARIMA200", "ADF", "AIC")
  
  set.seed(42)
  for (r in 1:reps){
    
    # Simulate time series of length n+1
    #   ARIMA(0,1,0) if phi1=1, ARIMA(1,0,0) otherwise
    if (phi1==1){
      Y <- arima.sim(n=n+1, model=list(order=c(0,1,0)))[-1]
    } else {
      Y <- arima.sim(n=n+1, model=list(ar=phi1))
    }
    
    # Keep aside Y[n+1] for forecast evaluation
    Yf <- Y[n+1]
    # Use Y1,...,Yn as the estimation sample
    Y <- Y[1:n]
    
    # Estimation of AR(2) model
    ARIMA200 <- lm(Y[t]~Y[t-1]+Y[t-2])
    
    # Estimation of ARIMA(1,1,0) model
    DY <- Y[t]-Y[t-1]
    DY1 <- Y[t-1]-Y[t-2]
    ARIMA110 <- lm(DY[t]~DY[t-1])
    
    
    # Compute one step ahead forecast errors
    b1 <- ARIMA110$coefficients
    ForecastErrors[r,1] <- Yf-(Y[n]+b1[1]+b1[2]*DY[length(DY)])
    
    b2 <- ARIMA200$coefficients
    ForecastErrors[r,2] <- Yf-(b2[1]+b2[2]*Y[n]+b2[3]*Y[n-1])
    
    
    # ADF test with drift, lags=1 (for p=2 model)
    # p>0.05 => H0:unit root not rejected, i.e. differencing indicated
    ur[r,1] <- 1*(punitroot(ur.df(Y, type="drift", lags=1)@teststat[1],
                            N=n, trend="c", statistic="t")>0.05)
    
    # If ADF test selects unit root use ARIMA110 forecast error:
    if (ur[r,1]==1){
      ForecastErrors[r,3] <- ForecastErrors[r,1]  
    } else { 
      # otherwise use ARIMA200 forecast error:
      ForecastErrors[r,3] <- ForecastErrors[r,2]  
    }
    
    # Use AIC to select between AR(2) and ARIMA(1,1,0)
    ur[r,2] <- 1*(AIC(ARIMA110)<AIC(ARIMA200))
    
    # If AIC selects unit root use ARIMA110 forecast error:
    if (ur[r,2]==1){
      ForecastErrors[r,4] <- ForecastErrors[r,1]  
    } else {
      # otherwise use ARIMA200 forecast error:
      ForecastErrors[r,4] <- ForecastErrors[r,2]  
    }
    
  }
  
  my_ur[i,] <- apply(ur, 2, mean)*100
  my_RMSE[i,] <- sqrt(apply(ForecastErrors^2, 2, mean))
  
}
```

Running the completed code produces the following results:
```{r}
print(my_ur)
```
<div style="height: 1em;"></div>
```{r}
print(round(my_RMSE,4))
```


## Results: Unit Root Detection {auto-animate="true"}
```{r}
print(t(my_ur))
```
<div class="horizontal-line"></div>
:::{.incremental}
- In this experiment the AIC demonstrates some tendency to find a unit root in a time series a little less than the ADF test, regardless of whether or not there is a unit root.
- When $\phi_{1}=1$ (unit root) the ADF test finds a unit root 95.8% of the time, consistent (approximately) with the 5% significance level of the test. The AIC finds a unit root 93.3% of the time. This does not appear to be a huge difference on the surface.
- When $\phi_{1}<1$ the difference between the ADF and AIC selection proportions becomes as high as about 7%. A noticeable but not huge difference. The implications of this difference are explored using forecast performance.
- <blue>Conclusion:</blue> Overall the AIC provides a reasonable way to distinguish between unit root or not in this experiment.
:::


## Results: RMSE {auto-animate="true"}
```{r}
print(round(my_RMSE,4))
```
<div class="horizontal-line"></div>

:::{.incremental}
- Many of the RMSEs are approximately the same. The biggest differences occur when is away from 1, i.e. $\phi_{1}=0.5,0.6$. 
  - In this case the ARIMA(1,1,0) model is clearly misspecified, and its forecasting performance suffers. 
  - However both ADF and AIC do a good job in selecting the correct model here.
  
- When $\phi_{1}=1$ the RMSEs are all roughly the same. This is interesting - if there is a unit root then it doesn’t matter what we do, at least for one-step-ahead forecasting.

- <blue>Overall the RMSEs for the ADF and AIC methods are practically the same.</blue> There is no benefit in making this decision in this experiment in using the specialised ADF, we could just as easily have used the AIC.

- Maybe interestingly, there is a “grey area’ around $\phi_{1}=0.95$ to $\phi_{1}=0.85$, where the selected models do worse than the ARIMA(2,0,0), although the RMSE differences are small.
:::


## Final Note
:::{.incremental}
- The most important thing is to understand the simulation approach.

- We are using simulation to replicate the frequentist statistical concept of a sampling distribution and hence to obtain experimental evidence on the performance of some tests and forecasts.

- <blue>Limitation: </blue>We have only tried a limited range of data generating processes and models and so the current results are only specific to that setting.
:::

## {data-background-color="#000F46"}
<div style="height: 85vh; display: flex; align-items: center; justify-content: center; color: white; font-size: 2em;">
Thank you for coming!
</div>


<script>
document.querySelectorAll("myimage").forEach(el => {
  const src = el.getAttribute("src");
  const width = el.getAttribute("width") || "100%";

  const img = document.createElement("img");
  img.src = src;
  img.style.width = width;
  img.style.maxWidth = "100%";  // prevent overflow
  el.style.display = "block";
  el.style.textAlign = "center";
  
  el.appendChild(img);
});
</script>